\large{Vipassana for Hackers} \\ \Huge{The Proposal} \\ \large\textit{Version 0.1}

### 2019-12-03 paper-three/paper-three.tex
proposal: finally done the intro, maybe
proposal: cite OPM MEG research paper, move method to method section

\begin{itemize}
  \item posture
  \item sleep
  \item digestion (``Make sure to pay attention to your poops!'' first course)
  \item diet ^
  \item schedule
  \item health (activation / motivation)
  \item ethics (activation / motivation)
  \item your children: a. knowing how to meditate, b. cross-legged posture
  \item emotion (i.e. anger)
  \item dealing with death
  \item mundane sphere / productivity (21 lessons, seinfeld)
  \item unlearning obsessive / repetitive thought, enhancing creativity
  \item controlling unbounded sexuality without repression (louis c.k. joke?)
  \item clarity: in thought, work, planning
  \item Die Standing Up
  \item \url{https://www.pnas.org/content/early/2019/10/18/1909959116}
\end{itemize}

@@

@@

@@

>>>>

\begin{itemize}
  \item posture
  \item sleep
  \item digestion (``Make sure to pay attention to your poops!'' first course)
  \item diet
  \item schedule
  \item health (activation / motivation)
  \item ethics (activation / motivation)
  \item your children: a. knowing how to meditate, b. cross-legged posture
  \item emotion (i.e. anger)
  \item dealing with death
  \item mundane sphere / productivity (21 lessons, seinfeld)
  \item unlearning obsessive / repetitive thought, enhancing creativity
  \item controlling unbounded sexuality without repression (louis c.k. joke?)
  \item clarity: in thought, work, planning
  \item Die Standing Up
  \item \url{https://www.pnas.org/content/early/2019/10/18/1909959116}
\end{itemize}

@@

\textbf{vedana-samosarana sabbe dhamma. ``Everything
that arises in the mind starts flowing with a sensation
on the body.''} \cite{goenka1999discourses}

@@

\bibitem{goenka1999discourses}
  Discourses on satipatthana sutta
%  @article{goenka1999discourses,
%  title={Discourses on satipatthana sutta},
%  author={Goenka, Satya Narayan},
%  journal={Igatpuri, Maharashtra, India: Vipassana Research Institute},
%  year={1999}
%}

@@

\bibitem{maruthai2016senior}
  Maruthai N., Nagendra R.P., Sasidharan A., Srikumar S., Datta K., Uchida S., Kutty B.M.
  \textit{Senior Vipassana Meditation practitioners exhibit distinct REM sleep
    organization from that of novice meditators and healthy controls.}
  International Review of Psychiatry, Volume 28(3), Pages 279-287, April 2016.
  \url{https://doi.org/10.3109/09540261.2016.1159949}

### 2019-12-03 the-proposal/the-proposal.tex
proposal: cite OPM MEG research paper, move method to method section

As the structure of studies conducted on Vipassana is inherently very difficult, due to
the nature of Vipassana meditation instruction, we must first examine which kinds of
studies are possible, which are not, and what will make possible studies worthwhile.

However, such a study suffers from the very nature of
the instruction and its target audience. On the lifelong scale of Vipassana
practice, the spectrum spans from laypeople practicing two hours per day to
renunciates (monks and nuns) who practice up to twenty-four hours per day once they
reach the stage where deep sleep no longer occurs due to the practice itself. It is
also worth noting that a junior renunciate will still have less \textit{experience}
than an senior lay meditator, which means that the only objective measure of
experience is total number of hours meditated. \cite{goleman2017altered}

Meditation research often has difficulty defining and capturing \textit{expertise},
however, and total hours of experience is still confounded by the rate in which those
hours of experience are accumulated (hours per day). We must capture both, as accurately as
possible. With a high hours-per-day rate and extended (lifelong) experience, the very
broad category of \textit{expert} meditators quickly requires upward of 100,000 hours
of practice, even if we limit practice hours to formal, sitting meditation and
estimate twelve (12) hours per day. If we take the much more modest rate of six (6)
hours per day, assume an adult can learn meditation from approximately 20 years of
age, and is of an age when they may participate in an extended study (say, 50-60
years of age), we arrive at roughly 60,000 hours of practice. If we use the oft-cited
``10,000 hours'' measurement for competence in a subject, we might reasonably
describe our rough spectrum of expertise as such:

As a consequence, even studies which claim to observe ``expert'' or ``long-term''
Vipassana meditators are often predominantly observing beginners (7.9-8.6 mean years
of experience with 2 hours of daily practice).

Chiesa also notes that it would be beneficial to capture data ``both from a clinical
and from a neuro-imaging point of view'', including functional magnetic resonance
imaging (fMRI) and electroencephalography (EEG), for both short-term outcomes
(altered states) and long-term outcomes (altered traits \cite{davidson1977role}) to
improve future Vipassana research. \cite{chiesa2010vipassana} New developments into
magnetoencephalography (MEG) and quantum MEG over the past decade may permit it as a
third neuro-imaging technique.

Because Vipassana does not permit any form of imagination, any attempt to analyze it
through philosophical phenomenology becomes a hindrance to actually practicing
Vipassana. \cite{patrik1994phenomenological} I have previously discussed this paradox
in Appendix A of \textit{Vipassana for Hackers, Paper One: Curious Mechanics}.
\cite{deobald2017curious}

@@

\section{Research Design and Methods}
\subsection{OVERVIEW}

>>>>

@@

\section{Research Design and Methods}

As the structure of studies conducted on Vipassana is inherently very difficult, due to
the strict nature of Vipassana meditation instruction, we must first examine which kinds of
studies are possible, which are not, and what will make possible studies worthwhile.

However, such a study suffers from the very nature of
the instruction and its target audience. On the lifelong scale of Vipassana
practice, the spectrum spans from a minimum of laypeople practicing two hours per day to
a maximum exemplified by renunciates (monks and nuns) who practice up to twenty-four
hours per day once they
reach the advanced stage where deep sleep no longer occurs. It is
also worth noting that a junior renunciate will still have less \textit{experience}
than a senior lay meditator, which means that the only objective measure of
experience is total number of hours meditated. \cite{goleman2017altered}

Meditation research often has difficulty defining and capturing \textit{expertise},
however, and total hours of experience is still confounded by the rate in which those
hours of experience are accumulated (hours per day). We must capture both, as accurately as
possible. Renunciates potentially have both a high hours-per-day rate and extended
(lifelong) duration of experience. The existence of renunciates forces the very
broad category of \textit{expert} meditators into the terrority of 100,000 hours
of practice, even if we limit practice hours to formal, sitting meditation and
estimate twelve (12) hours per day. If we take the much more modest rate of six (6)
hours per day, assume an adult renunciate can learn Vipassana, at the earliest, from
approximately 20 years of
age, and is now of an age when they may participate in an extended study (say, 50-60
years of age), we arrive at roughly 60,000 hours of practice. If we use the oft-cited
``10,000 hours'' measurement for competence in a subject, we might reasonably
describe our rough spectrum of expertise as such:

As a consequence, even studies which claim to observe ``expert'' or ``long-term''
Vipassana meditators are often predominantly observing beginners (7.9-8.6 mean years
of experience with 2 hours of daily practice). \cite{chiesa2010vipassana}

Chiesa also notes that it would be beneficial to capture data ``both from a clinical
and from a neuro-imaging point of view'', including functional magnetic resonance
imaging (fMRI) and electroencephalography (EEG), for both short-term outcomes
(altered states) and long-term outcomes (altered traits \cite{davidson1977role}) to
improve future Vipassana research. \cite{chiesa2010vipassana}
Optically-pumped magnetometer (OPM)-based magnetoencephalography (MEG) performed using a
mobile helmet capable of operating at room temperature \cite{boto2017new}, in development since Chiesa's
paper was published, may permit it as a third neuro-imaging technique. Even older MEG
technology based on superconducting quantum interference devices (SQUIDs) could
potentially be used. The restriction of SQUID-based MEG is that subjects must remain
extremely still, but this is all but a requirement of serious Vipassana meditation
anyway.

Because Vipassana does not permit any form of imagination, any attempt to analyze it
through philosophical phenomenology becomes a hindrance to actually practicing
Vipassana. \cite{patrik1994phenomenological} The two are mututally exclusive, as is
any attempt to contemplate Vipassana, the phenomenology of the mind, or the
technique, during the actual practice of Vipassana. I have previously discussed this
apparent paradox in Appendix A of \textit{Vipassana for Hackers, Paper One: Curious Mechanics}.
\cite{deobald2017curious}
